---
title: "Introduction to R Notebooks"
output: html_notebook
---

### Objectives

By the time you work through this document you will be able to...

 - Write and execute R code within an R chunk, both line-by-line and as a group
 - Follow instructions within an R Notebook
 - Use the Preview button to create an HTML document of your work.
 
# Introduction and Installing Packages 

Welcome to your first R Notebook.  This document seamlessly integrates text, R code, R output, and R figures.  Most importantly, it makes it possible for you to complete your entire lab within R Studio without having to toggle between Canvas, a pdf, and a separate document for your responses. This document can be converted to an HTML that combines your typed responses, as well as all your R output and all your figures. 

You will work your way through this document to learn the basics of Notebooks.  Watch for the **TASK**s, which indicate that you need to perfom a calculation in R or answer a question.  Your first task is below this sentence.

**TASK 0**  Make this pane occupy the whole screen.  That means, reduce the console, the environment, and the plots tabs. (Use the button in the upper right corner of this pane.)

**TASK 1** . Above this pane, but below the tabs with the file names of the documents you have open, you should see a button that says simply "Preview".  Click on Preview. When you do, R will likely ask you to install some packages. Say yes.  It may take five to ten minutes. 

Once that is complete, you will want to specifically install the packages of "mosaic" and "ggformula". These will be utilized in all R Lab assignments in the course. Again, this may take a few minutes, depending on the machine. To install these packages, click the green arrow button on the right side of the R code chunk.

```{r, echo = F}
install.packages("mosaic",configure.args=c("--with-gdal-config=/usr/bin"))
install.packages("ggformula")

```

After installing the packages, you will need to load these packages. Run the chunk of R code below. 

```{r, echo = F}
library(mosaic, warn.conflicts = FALSE) 
library(ggformula, warn.conflicts = FALSE)
```


Now hit the Preview button. You should have seen an HTML pop up of this assignment, complete with all the directions and questions.  You can close that window now.

### R Chunks

There are two different kinds of work you can do in this Notebook.  Normal text, which is what you read now, and R code.  You can code in R using *R Chunks*.  Your first R chunk is below:

```{r, echo = F}
print("Hello World")
```

To run a line of code within an R chunk, place your cursor on that line and press Ctrl+Enter on a windows machine, or Command+Enter on a mac.  

**TASK 2** Use Ctrl+Enter or Cmd+Enter to run the line of code below:

```{r, echo = F}
print("Hello World")
```

**TASK 3** Now create an HTML using Preview.  How many times do you see "Hello World" in your HTML? Type your answer below next to **Response:**

    **Response:*
    
**TASK 4** R contains many built-in datasets. The dataset "cars" contains the speed and stopping distance for a set of cars. The code below will allow you to view numeric and graphical summaries of this dataset.

Run all the code at once in the chunk below by clicking on the green arrow on the right side of the code chunk. The code below will generate several plots and summary statistics. The specifics of how this code is written will be covered in your first R Lab 1.1.

```{r, echo = F}
favstats(~speed, data = cars)
gf_point(speed ~dist, data = cars)
gf_histogram(~speed,data = cars)
```

Notice that when you run multiple lines at once you can click through the figures and output.  When you create the HTML, all figures and output will be included.


**TASK 4** Create an HTML using Preview again.  This time what is different from your HTML in TASK 3?

    **Response:**
    
# Closing Notes

You now know the basics of using an R Notebook.  From now on your labs will be completed as R Notebooks, and you will submit your work as an HTML in Canvas.  It is important to note that your R output (figures and calculations) will only be included in the HTML if you ran the code in the Notebook.  If you write the code but don't run it, your work will not be included in the HTML.

To create a new R chunk you can type use the shortcut Ctrl+Alt+i or Cmnd+Alt+i.

