---
title: 'Final Exam R Shell'
output: html_notebook
---


### introduction:

You can use this document as a place to read in your data and write your code for the final exam.  You will need to write all your own code, but you can reference the labs that you already completed.
 
Reminder: to create a vector of values in R use code that looks like this:

myvector <- c(1,2,3,4)


```{r, echo = F, message = F}
##  Run this piece of code to load the packages 
# Clear Workspace
rm(list = ls()) 

# load packages we typically use for this class.
library(mosaic, warn.conflicts = FALSE) 
library(ggformula, warn.conflicts = FALSE)
```
# Questions 1-10

You can use the R chunk below to perform any calculations you may want.  There are no datasets to load for these questions.

```{r}
pbinom(19,100,.15)
X = c(1.8,4.7,.7,1.43,.66)
gf_qq(~X) %>% gf_qqline()
gf_histogram(~X)
gf_boxplot(~X)
pnorm(.5,2,1) - pnorm(-1,2,1)
```

# Question 11

Use the R chunk below to load the data and perform any analysis needed.

```{r}
# run this code chunk to load the student data
mydata11 <- read.csv(file.choose())
t.test(MidtermScore ~ Section,conf.level = .95,data = mydata11)$conf.int
favstats(MidtermScore ~ Section,data = mydata11)
```

# Questions 12-17

Use the code chunk below to perform any calculations for these questions that do not use data.

```{r}

```

# Question 18

Use the code chunk below to read in the data and perform any needed calculations.

```{r}
# run this code chunk to load the student data
mydata18 <- read.csv(file.choose())
favstats(~BloodPressure,data = mydata18)
t.test(~BloodPressure,alternative = "greater",mu = 120, data = mydata18)

```

# Questions 19-20

Use the code chunk below to perform any calculations for these questions that do not use data.

```{r}
X = c(23, 36, 21, 18)
chisq.test(X)
```

# Question 21

Use the code chunk below to read in the data and perform any needed calculations.

```{r}
# run this code chunk to load the student data
mydata21 <- read.csv(file.choose())
mymodel <- lm(MPG~WT,data = mydata21)
summary(mymodel)
```

# Questions 22-23

Use the code chunk below to perform any calculations for these questions that do not use data.

```{r}

```

# Question 24

Use the code chunk below to read in the data and perform any needed calculations.

```{r}
# run this code chunk to load the student data
mydata24 <- read.csv(file.choose())
favstats(Ozone ~ Month,data = mydata24)
```


# Remaining Questions

Use the code chunk below to perform any calculations for these questions that do not use data.

```{r}

```