---
title: "Lab 2-2: Binomial and Exponential Distributions"
output: html_notebook
---

### Objectives:

To use the built-in function in R to compute probabilities using specific probability distributions.

# Part 1: Binomial Distribution

R has the ability to calculate probabilities exactly from a wide variety of distributions. For Binomial distributions:

  - dbinom(x, size = n, prob = p) calculates P(X=x) for X ~ binomial(n,p)
  - pbinom(x,size = n, prob = p) calculates $P(X ≤ x)$ for X ~ binomial(n,p)
  - qbinom(pth,size = n, prob = p) calculates the pth quantile for X ~  binomial(n,p)
  - rbinom(N,size = n,rob = p) simulates N observations from a binomial(n,p)
  


 Assume that X is the number of heads obtained when flipping a weighted coin 100 times, where the probability the coin lands on heads for any particular flip is 0.09.

**TASK 1.1**  Find P(X=9) using R.

**Response:** P(X=9) = 0.138

```{r, echo = F}
dbinom(9,100,0.09)
```

**TASK 1.2** Show how the value in 1.1 is obtained using the probability mass function for X.

**Response:** $\frac{100!}{(9!(100-9)!)}(0.09^9)(1-0.09)^{100-9} = 0.138$

**TASK 1.3** Find $P(X ≤ 9)$ using R.

**Response:** $P(X ≤ 9)$ = 0.588

```{r, echo = F}
pbinom(9,100,0.09)
```

**TASK 1.4** Find $P(X ≤  20)$ using R.

**Response:**  $P(X ≤  20)$ = 0.9998
```{r, echo = F}
pbinom(20,100,0.09)
```

**TASK 1.5** Explain why your calculations in 1.4 and 1.5 would be cumbersome to do by hand.

**Response:**  They would involve sums with lots of combinations.

**TASK 1.6**  Find $P(5 ≤ X ≤ 15)$ using R.  Hint - you'll need to use pbinom() twice.

**Response:** $P(5 ≤ X ≤ 15)$ = 0.936
```{r, echo = F}
pbinom(15,100,0.09) - pbinom(4,100,0.09)
```


Now assume that the probability of a coin landing directly on its side 0.005.  Let Y = \# side lands when flipping the coin 1000 times.

**TASK 1.7**  Find $P(3 ≤ Y ≤ 6)$ using R.

**Response:** $P(3 ≤ Y ≤ 6)$ = 0.639
```{r, echo = F}
pbinom(6,1000,0.005) - pbinom(2,1000,0.005)
```

**TASK 1.8** Use the fact that Y ~ Binom(1000, 0.005) to find the expected value of Y.

**Response:**  $\mu = 5$

**TASK 1.9** Find the standard deviation of Y.

**Response:** $\sigma = 2.230$

**TASK 1.10** Use your responses for 1.8 and 1.9, to explain whether the value of $P(3 ≤ Y ≤ 6)$ that you calculated in 1.7 seems reasonable.

**Response:** Because $P(3 ≤ Y ≤ 6)$ is contained by the mean plus or minus the standard deviation it makes sense that the probability would be large

# Part 2: Exponential

A well-known continuous distribution is the Exponential Distribution.  A random variable that follows this distribution has probability density function
\[f(x) =\lambda e^{-\lambda x},\ \ \ x>0\]

The parameter $\lambda$ is often referred to as the rate parameter, and a common use of this distribution is to model waiting time until an event occurs.  

Again, R has the ability to calculate probabilities exactly an exponential distribution:

  - pexp(x, rate = lambda) calculates $P(X ≤ x)$ for X ~ Exp(lambda)
  - qexp(pth, rate = lambda) calculates the pth quantile for X ~ Exp(lambda)
  - rexp(N, rate = lambda) simulates N observations from Exp(lambda)


Assume the average time  X that it takes for a delayed plan to depart is half an hour, and is modeled as an Exponential random variable where X is in hours and $\lambda = 2$.

**TASK 2.1**  Use R to find the probability that the delayed plane waits less than 20 minutes before departing.  Hint - 20 minutes is 1/3 hour.

**Response:** $P(X ≤ 1/3)$ = 0.487
```{r, echo = F}
pexp(1/3, rate = 2)
```

**TASK 2.2** If you were to write the integral to compute the probability from part 2.1, what would be the limits of integration and the integrand?

**Response:**
   - integrand: $2 e^{-2 x}$
   - lower limit: 0
   - upper limit: 1/3

**TASK 2.3** Based on the probability that you calculated in 2.1, try different values for the median wait time for the plan to depart, for example try pexp(1/4,rate = 2) and see if it's closer to the median than 1/3.  Try several values and state your closest value here as your response:

**Response:**  0.347 hours
```{r}
pexp(0.3465736, rate = 2)
```

**TASK 2.4** Use R find P(1/3 ≤ X ≤ 1/2).

**Response:** $P(1/3 ≤ X ≤ 1/2)$ = 0.146
```{r, echo = F}
pexp(1/2, rate = 2) - pexp(1/3, rate = 2)
```

**TASK 2.5**  What is the probability that you wait exactly 35 minutes to depart?

**Response:**  $P(X = 35/60)$ = 0


